An Implementation of the Active Contours without Edges model and the Logic Framework for Active contours on multi-channel images

### 2010-12-13 cs870/project/Report/Paper.tex

\usepackage{cvpr}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=true]{hyperref}

@@

\begin{abstract}
In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It differs
from other active contour modesl in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
contour to evolve simultaneously on multiple channels. It also offers a set of logic operations that can be applied to those channels (union, intersection,
complement, and their combinations). We evaluate our implementation based on several criteria, such as: independence of initial curve position, successful
detection of holes, performance with blurred and noisy images, and response to scaling parameters.
\end{abstract}

\section{Introduction}
Image segmentation or boundary detection is a very important problem in the area of Image Processing, and has received a lot of attention in the past. The
classical Active Contour model (or Snakes) proposed by Kass et al. \cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
(E.g.~\cite{caselles1997geodesic}. Most of these models used the level set formulation for propagating fronts to evolve the curve. However, these models highly
depended on curvature motion (motion defined by the gradient of the curve) which led to poor performance in smoothed edges. To overcome these limitations, Chan
and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model \cite{mumford1989optimal} to divide the image
into two regions: one inside the propagating curve, and one outside. The curve is at the boundary of the object if there is no difference in intensities inside
the curve as well as outside the curve.

The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) was very successful in
detecting objects even in noisy or blurry images. It could also detect holes in objects which was usually a limitation in previous models. As an extension to
this work, Sandberg and Chan~\cite{sandberg2005logic} propose a logic framework (referred to as the Sandberg-Chan model throughout the rest of this
paper) that  performs logical operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
through bitwise operations (CITE) or they would apply logic operations to the different images then segment the resulting image (CITE). The first approach is
very costly, and the second approach requires a lot of prior knowledge about the intensities of each image. To overcome these drawbacks, the Sandberg-Chan
model is based on the idea of fitting a single contour to the object on all channels according to the logic operator, and based on regions.

@@

(REST OF EQUATION HERE)

@@

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{logicop.png}
\caption{Region Based View Union}
\label{fig:logic-op}
\end{figure*}

@@

with the boundary condition

$\frac{\delta(\phi)}{|\bigtriangledown\phi|}\frac{\partial\phi}{\partial n^{\rightarrow}} = 0$

@@

@@

\begin{figure*}[t]
\centering
\includegraphics[width=11cm]{cv_eg1.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
\label{fig:cv_eg1}
\end{figure*}

@@

@@

To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same brain image
used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
positions. The first one is shown in Figure~\ref{fig:cv_eg3} where the initial contour only partially overlaps with the object. The second one is shown in
Figure~\ref{fig:cv_eg4} where the initial contour does not overlap the object in any part. Visually, the obtained segmentation was the same with a slightly
higher number of iterations and CPU time. In all three cases, the brain was correctly segmented. We compared the area segmented as the brain in each case to
make sure the same segmentation was obtained, and in each case, we got the same area of 0.5268. Note, however, that when part of the curve lies outside the
object, the definition of inside and outside changes since the curve becomes an open curve at one point. This should not make a difference as long as the
contour lies correctly at the boundaries of the object.

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg3.png}
\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
\label{fig:cv_eg3}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg4.png}
\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
\label{fig:cv_eg4}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg6.png}
\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
\label{fig:cv_eg6}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=10cm]{cv_eg7.png}
\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
\label{fig:cv_eg7}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg9.png}
\caption{Ability to detect contours in blurry images.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 255 x 255, $\phi_{0}(x,y) = - \sqrt{(x - 127.5)^2 + (y - 127.5)^2)} +
63.75$, $\mu =0.01$, no reinitialization, cpu = 5.41 s, took 15 iterations.}
\label{fig:cv_eg9}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg11.png}
\caption{Same image as Figure~\ref{fig:cv_eg10}, but with 'salt \& pepper' noise. Noise was still detected despite increasing $\mu$.  Top: the evolving curve
(in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
75$, $\mu = 5$, no reinitialization, cpu = 12.27 s, took 7 iterations.}
\label{fig:cv_eg11}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{threech.png}
\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
5. For intersection then union: cpu = 1.24 and iterations = 4.}
\label{fig:threech}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{cv_eg12.png}
\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
\label{fig:cv_eg12}
\end{figure}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{donutlogic.png}
\caption{Performing union and intersection on images with holes. Size = 300 * 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
75$, no reinitialization. For union: cpu = 8.63 sec and iterations = 10. For intersection: cpu = 7.68 sec and iterations = 10.}
\label{fig:donutlogic}
\end{figure*}

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{family.png}
\caption{Combining iterations = 7. union 4.01 s intersection 4.12 s}
\label{fig:donutlogic}
\end{figure}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{flowers.png}
\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
size = cpu = 1.73 sec, iterations = 7.}
\label{fig:flowers}
\end{figure*}

@@

@@

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{sc_position.png}
\caption{Same example shown in Figure~\ref{fig:sc_unionsimple}, but with different initial curve position. Initial curve does not overlap with either objects.
Same segmentation is obtained. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 4.43s, 4
iterations.}
\label{fig:sc_position}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{sc_noisy.png}
\caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
50$, no reinitialization. Cpu = 3.17 sec and iterations = 5 when $\lambda = 255*255$. Cpu = 146.46 sec and iterations = 125 when $\lambda =
125$.}
\label{fig:sc_noisy}
\end{figure}

@@

\subsection{Varying $\mu$}
$\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure [\ref{fig:cvmu}] shows the
segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{cvmu.png}
\caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing
purposes. Top: original image with noise. Bottom: segmentation for various values of $\mu$.}
\label{fig:cvmu}
\end{figure*}

@@

\begin{equation}
\label{eqn:complement}
\begin{align*}
{z_{i}}^{in^{'}} = 1 - {z_{i}}^{in}, \\
{z_{i}}^{out^{'}} = 1 - {z_{i}}^{out}.
\end{align*}
\end{equation}

Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation
[\ref{eqn:complement}], and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure [\ref{fig:sccomp}] shows this scenrio. We tried to
debug the reasons for this problem for three days.

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{sccomp.png}
\caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
of $A_1 \cap A_2$. Bottom: correct and incorrect segmentation respectively}
\label{fig:sccomp}
\end{figure*}

\subsection{Defining the initial contour ($\phi_0$)}
Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{phinote.png}
\caption{Assumptions about the way $\phi_0$ is defined.}
\label{fig:phinote}
\end{figure}

@@

Although this definition of $\phi_0$ is compatible with most of the images we used to test our implementation, it did fail to correctly segment the objects in
some images. In fact, the results were the opposite of the used logic operator (i.e. the output is the intersection of the channels, while it should be the
union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
circle to initialize $\phi_0$ (Equation [\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
of the input channels. Figure [\ref{fig:phidef}] shows the final segmentation using both definitions for $\phi_0$.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{phidef.png}
\caption{}
\label{fig:phidef}
\end{figure*}

\subsection{Mistyped Equations}
As previously mentiond in Section [\ref{sec:sandberg-chan}], there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
(equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
typos consumed over a week of our time until we were able to find out what are the correct equations in both cases. We could have spent this time overcoming
other challenges that we faced while implementing both models.

@@

\bibliographystyle{IEEEtran}
\bibliography{references}

>>>>

\usepackage{cvpr}
\usepackage{times}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi

@@

\begin{abstract}
In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It
differs
from other active contour models in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
contour to evolve simultaneously on multiple channels. It also offers a set of logic operations that can be applied to those channels (union, intersection,
complement, and their combinations). We evaluate our implementation based on several criteria, such as: independence of initial curve position, successful
detection of holes, performance with blurred and noisy images, and response to scaling parameters.
\end{abstract}

\section{Introduction}
Image segmentation or boundary detection is a very important problem in the area of Image Processing, and has received a lot of attention in the past. The
classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
(E.g.~\cite{caselles1997geodesic}. Most of these models used the level set formulation for propagating fronts to evolve the curve. However, these models highly
depended on curvature motion (motion defined by the gradient of the curve) which led to poor performance in smoothed edges. To overcome these limitations, Chan
and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model \cite{mumford1989optimal} to divide the image
into two regions: one inside the propagating curve, and one outside. The curve is at the boundary of the object if there is no difference in intensities inside
the curve as well as outside the curve.

The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) was very successful in
detecting objects even in noisy or blurry images. It could also detect holes in objects which was usually a limitation in previous models. As an extension to
this work, Sandberg and Chan~\cite{sandberg2005logic} propose a logic framework (referred to as the Sandberg-Chan model throughout the rest of this
paper) that  performs logical operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
through bitwise operations (E.g.~\cite{sapiro1997color, zhu2002region}) or they would apply logic operations to the different images then segment the resulting
image. The first approach is
very costly, and the second approach requires a lot of prior knowledge about the intensities of each image. To overcome these drawbacks, the Sandberg-Chan
model is based on the idea of fitting a single contour to the object on all channels according to the logic operator, and based on regions.

@@

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{logicop.png}
\caption{Region Based View Union}
\label{fig:logic-op}
\end{figure*}

@@

@@

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{cv_eg1.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
\label{fig:cv_eg1}
\end{figure}

@@

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{cv_eg3.png}
\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
\label{fig:cv_eg3}
\end{figure*}

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{cv_eg4.png}
\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
\label{fig:cv_eg4}
\end{figure*}

@@

To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the lesame brain image
used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
positions. The first one is shown in Figure~\ref{fig:cv_eg3} where the initial contour only partially overlaps with the object. The second one is shown in
Figure~\ref{fig:cv_eg4} where the initial contour does not overlap the object in any part. Visually, the obtained segmentation was the same with a slightly
higher number of iterations and CPU time. In all three cases, the brain was correctly segmented. We compared the area segmented as the brain in each case to
make sure the same segmentation was obtained, and in each case, we got the same area of 0.5268. Note, however, that when part of the curve lies outside the
object, the definition of inside and outside changes since the curve becomes an open curve at one point. This should not make a difference as long as the
contour lies correctly at the boundaries of the object.

\begin{figure*}[t!h!]
\centering
\includegraphics[width=12cm]{cv_eg6.png}
\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
\label{fig:cv_eg6}
\end{figure*}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{cv_eg7.png}
\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
\label{fig:cv_eg7}
\end{figure}

@@

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{cv_eg9.png}
\caption{Ability to detect contours in blurry images.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 255 x 255, $\phi_{0}(x,y) = - \sqrt{(x - 127.5)^2 + (y - 127.5)^2)} +
63.75$, $\mu =0.01$, no reinitialization, cpu = 5.41 s, took 15 iterations.}
\label{fig:cv_eg9}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{cv_eg11.png}
\caption{Same image as Figure~\ref{fig:cv_eg10}, but with 'salt \& pepper' noise. Noise was still detected despite increasing $\mu$.  Top: the evolving curve
(in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
75$, $\mu = 5$, no reinitialization, cpu = 12.27 s, took 7 iterations.}
\label{fig:cv_eg11}
\end{figure*}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=0.5\textwidth]{cv_eg12.png}
\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
\label{fig:cv_eg12}
\end{figure}

@@

@@

\begin{figure*}[t!h!]
\centering
\includegraphics[width=12cm]{donutlogic.png}
\caption{Performing union and intersection on images with holes. Size = 300 * 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
75$, no reinitialization. For union: cpu = 8.63 sec and iterations = 10. For intersection: cpu = 7.68 sec and iterations = 10.}
\label{fig:donutlogic}
\end{figure*}

\begin{figure}[t!h!]
\centering
\includegraphics[width=0.3\textwidth]{family.png}
\caption{Combining missing information from two different channels. Size = 114 * 114. $\phi_{0}(x,y) = - \sqrt{(x - 57)^2 + (y - 57)^2)} +
28.5$, no reinitialization. Union: cpu = 4.01 s and iterations = 7. Intersection: cpu = 4.12 s and iterations = 7.}
\label{fig:family}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.3\textwidth]{sc_position.png}
\caption{Same example shown in Figure~\ref{fig:sc_unionsimple}, but with different initial curve position. Initial curve does not overlap with either objects.
Same segmentation is obtained. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 4.43s, 4
iterations.}
\label{fig:sc_position}
\end{figure}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{flowers.png}
\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
size = cpu = 1.73 sec, iterations = 7.}
\label{fig:flowers}
\end{figure*}

\begin{figure}[t!]
\centering
\includegraphics[width=0.3\textwidth]{threech.png}
\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
5. For intersection then union: cpu = 1.24 and iterations = 4.}
\label{fig:threech}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{sc_noisy.png}
\caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
50$, no reinitialization. Cpu = 3.17 sec and iterations = 5 when $\lambda = 255*255$. Cpu = 146.46 sec and iterations = 125 when $\lambda =
125$.}
\label{fig:sc_noisy}
\end{figure}

@@

\subsection{Varying mu}
$\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure~\ref{fig:cvmu} shows the
segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{cvmu.png}
\caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing
purposes. Top: original image with noise. Bottom: segmentation for various values of $\mu$.}
\label{fig:cvmu}
\end{figure}

@@

\begin{equation}
\label{eqn:complement}
%\begin{align*}
{z_{i}}^{in^{'}} = 1 - {z_{i}}^{in}, \\
{z_{i}}^{out^{'}} = 1 - {z_{i}}^{out}.
%\end{align*}
\end{equation}

Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation~\ref{eqn:complement},
and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure~\ref{fig:sccomp} shows this scenrio. We tried to
debug the reasons for this problem for three days.

\begin{figure}[t]
\centering
\includegraphics[width=0.25\textwidth]{sccomp.png}
\caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
of $A_1 \cap A_2$. Bottom: correct and incorrect segmentation respectively}
\label{fig:sccomp}
\end{figure}

\subsection{Defining the initial contour phi0}
Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure~\ref{fig:phinote} illustrates those assumptions.

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{phinote.png}
\caption{Assumptions about the way $\phi_0$ is defined.}
\label{fig:phinote}
\end{figure}

@@

Although this definition of $\phi_0$ is compatible with most of the images we used to test our implementation, it did fail to correctly segment the objects in
some images. In fact, the results were the opposite of the used logic operator (i.e. the output is the intersection of the channels, while it should be the
union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
circle to initialize $\phi_0$ (Equation~\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
of the input channels. Figure~\ref{fig:phidef} shows the final segmentation using both definitions for $\phi_0$.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.2\textwidth]{phidef.png}
\caption{}
\label{fig:phidef}
\end{figure}

\subsection{Mistyped Equations}
As previously mentiond in Section~\ref{sec:sandberg-chan}, there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
(equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
typos consumed over a week of our time until we were able to find out what are the correct equations in both cases. We could have spent this time overcoming
other challenges that we faced while implementing both models.

@@

\bibliography{references}
\bibliographystyle{ieee}
